---
title: "Wolves"
author: "Kellie Liu"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
    keep.md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the Libraries  
```{r}
library(tidyverse)
library(shiny)
library(shinydashboard)
require(janitor)
library(palmerpenguins)
```

## Practice
Let's try to build an app that allows users to explore sexual composition of wolf packs by `pop`. These data are from: Brandell, Ellen E (2021), Serological dataset and R code for: Patterns and processes of pathogen exposure in gray wolves across North America, Dryad, [Dataset](https://doi.org/10.5061/dryad.5hqbzkh51)  

1. Start by loading and exploring the data.
```{r}
wolves <- read_csv("data/wolves_data/wolves_dataset.csv")
```

```{r}
names(wolves)
```

```{r}
wolves
```

2. For the app, here is a faceted version of what we are looking for. We want the reactive part to be `pop` and you should use `shinydashboard`.
```{r}
wolves %>% 
  filter(sex!="NA") %>% 
  ggplot(aes(x=sex, fill=sex))+
  geom_bar()+
  facet_wrap(~pop)
```

```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Wolves"),
  dashboardSidebar(disable=T),
  dashboardBody(
    selectInput("pop", "Select Pop", 
                choices = c("AK.PEN", "BAN.JAS", "BC", "DENALI", 
                            "ELLES", "GTNP", "INT.AK", "MEXICAN", 
                            "MI", "MT", "N.NWT", "ONT", "SE.AK", 
                            "SNF", "SS.NWT", "YNP", "YUCH"),
                selected = "AK.PEN"),
    plotOutput("plot", width = "600px", height = "500px")
  )
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    wolves %>% 
      filter(sex != "NA", pop == input$pop) %>% 
      ggplot(aes(x = sex, fill = sex)) +
      geom_bar() +
      facet_wrap(~pop)
  })
}

shinyApp(ui, server)

```
