---
title: "lab5_warmup"
output: html_document
date: "2023-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
### 1. Load the tidyverse
```{r}
library("tidyverse")
```

### 2. Load the bison data
```{r}
bison <- read_csv("data/bison.csv")
```

### 3. Evaluate the bison data
```{r}
glimpse(bison)
```

### 4. Pull out code, sex, weight, year of birth
```{r}
bison_new <- select(bison, "animal_code", "animal_sex", "animal_weight", "animal_yob")
```

### 5. Pull out the animals born between 1980-1990
```{r}
bison_yob <- filter(bison_new, animal_yob>=1980 & animal_yob<=1990)
```

### 6. How many males and females are represented in the data between 1980-1990?
```{r}
table(bison_yob$animal_sex)
```

### 7. In the 1980-1990 data, are females or males larger on average?
```{r}
bison_m <- filter(bison_yob, animal_sex=="M")
bison_f <- filter(bison_yob, animal_sex=="F")
```

### 8. Means
```{r}
mean(bison_m$animal_weight) #males
```

```{r}
mean(bison_f$animal_weight) #females
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
